### 2015-09-11 BPLE/Manuscript.tex
first commit. empty repo

% an empty .tex file. Eventually may transwer .odt in here